---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r,echo=TRUE}
if(!file.exists("activity.csv")) {
        tempfile <- tempfile()
        download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",destfile = tempfile)
        unzip(tempfile)
        unlink(tempfile)
}
activity <- read.csv("activity.csv") #Loading data
summary(activity)
str(activity)

```
## What is mean total number of steps taken per day?

#Total number of steps taken per day
```{r,echo=TRUE}
activity_steps_day <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
```
#Histogram of the total number of steps taken each day
```{r,echo=TRUE}
hist(activity_steps_day$steps, xlab = "Steps per Day", main = "Total number of steps taken per day", col = "wheat")
```
# Mean and median of the total number of steps taken per day
```{r,echo=TRUE}
mean_steps <- mean(activity_steps_day$steps)
median_steps <- median(activity_steps_day$steps)

mean_steps <- format(mean_steps,digits=1)
median_steps <- format(median_steps,digits=1)
```
### Mean steps per day: 10766
### Median steps per day: 10765

## What is the average daily activity pattern?

#Time series plot
```{r,echo=TRUE}
activity_steps_mean <- aggregate(steps ~ interval, data = activity, FUN = mean, na.rm = TRUE)
plot(activity_steps_mean$interval, activity_steps_mean$steps, type = "l", col = "tan3", xlab = "Intervals", ylab = "Total steps per interval", main = "Number of steps per interval (averaged) (NA removed)")
```
#The 5-minute interval that, on average, contains the maximum number of steps
```{r,echo=TRUE}
max_steps <-max(activity_steps_mean$steps)
max_interval <- activity_steps_mean$interval[which(activity_steps_mean$steps == max_steps)]
max_steps <- round(max_steps, digits = 2)
```
### The highest number of steps for a 5 minutes interval is 206.17


## Imputing missing values

#1.Calculate total number of missing values in the dataset
```{r,echo=TRUE}
sum(is.na(activity))#Total Number of missing values
```
### The total number of missing values is 2304.

#2.Devise a strategy for filling in all of the missing values in the dataset
```{r,echo=TRUE}
missing_values <- subset(activity, is.na(steps))#subset general dataset with only missing values 
par(mfrow = c(2,1), mar = c(2, 2, 1, 1))
hist(missing_values$interval, main="NAs repartition per interval")
hist(as.numeric(missing_values$date), main = "NAs repartion per date", breaks = 61)#Plot of repartition by date or by intervals
```
#3. Create new dataset with the missing data filled in
```{r,echo=TRUE}
MeanStepsPerInterval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)#Mean of steps per interval
activity_NAs <- activity[is.na(activity$steps),]
activity_non_NAs <- activity[!is.na(activity$steps),]# Cut the 'activity' dataset in 2 parts
activity_NAs$steps <- as.factor(activity_NAs$interval)
levels(activity_NAs$steps) <- MeanStepsPerInterval# Replace missing values in activity_NAs
levels(activity_NAs$steps) <- round(as.numeric(levels(activity_NAs$steps)))
activity_NAs$steps <- as.integer(as.vector(activity_NAs$steps))#Change the vector back as integer 
imputed_activity <- rbind(activity_NAs, activity_non_NAs)#Merge the two datasets together
```
#4. A histogram of the total number of steps taken each day
```{r,echo=TRUE}
#Plotting parameters to place previous histogram and new one next to each other
par(mfrow = c(1,2))
#Plot again the histogram from the first part of the assignment
activity_steps_day <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
hist(activity_steps_day$steps, xlab = "Steps per Day", main = "NAs REMOVED - Total steps/day", col = "wheat")
#Plot new histogram, with imputed missing values
imp_activity_steps_day <- aggregate(steps ~ date, data = imputed_activity, FUN = sum, na.rm = TRUE)
hist(imp_activity_steps_day$steps, xlab = "Steps per Day", main = "NAs IMPUTED - Total steps/day", col = "wheat")
```

```{r,echo=TRUE}
imp_mean_steps <- mean(imp_activity_steps_day$steps)
imp_median_steps <- median(imp_activity_steps_day$steps)
#we set a normal number format to display the results
imp_mean_steps <- format(imp_mean_steps,digits=1)
imp_median_steps <- format(imp_median_steps,digits=1)
#store the results in a dataframe
results_mean_median <- data.frame(c(mean_steps, median_steps), c(imp_mean_steps, imp_median_steps))
colnames(results_mean_median) <- c("NA removed", "Imputed NA values")
rownames(results_mean_median) <- c("mean", "median")
```
### Imputed mean steps per day: 10766
### Imputed median steps per day: 10762

## Are there differences in activity patterns between weekdays and weekends?

#1. Create a new factor variable
```{r,echo=TRUE}
imputed_activity$dayType <- ifelse(weekdays(as.Date(imputed_activity$date)) == "Samstag" | weekdays(as.Date(imputed_activity$date)) == "Sonntag", "weekend", "weekday")
imputed_activity$dayType <- factor(imputed_activity$dayType)#Transform into factor
```
#2.Panel plot containing time series plot
```{r,echo=TRUE}
steps_interval_dayType <- aggregate(steps ~ interval + dayType, data = imputed_activity, FUN = mean)
head(steps_interval_dayType)
names(steps_interval_dayType) <- c("interval", "day_type", "mean_steps")
```
```{r,echo=TRUE}
library(ggplot2)
plot <- ggplot(steps_interval_dayType, aes(interval, mean_steps))
plot + geom_line(color = "tan3") + facet_grid(day_type~.) + labs(x = "Intervals", y = "Average Steps", title = "Activity Patterns")
```
